Improved Quasi-Static Methods in Rattlesnake

### 2016-06-14 doc/NEAMS_Report/NEAMS_Report.tex
updated abstract

Because of the recent interest in reactor transient modeling and the restart of the Transient Reactor (TREAT) Facility, there has been a need for more efficient, robust methods in computation frameworks.  This is the impetus of implementing the Improved Quasi-Static method (IQS) in the Rattlesnake framework.  IQS has implemented by factorizing flux into time-dependent amplitude and spacial- and weakly time-dependent shape.  The shape evaluation is very similar to a flux solve and is computed at large (macro) time steps.  While the amplitude evaluation is a PRKE solve where the parameters are dependent on the shape and is computed at small (micro) time steps.  IQS has been tested with a custom one-dimensional example, the TWIGL ramp benchmark, and LRA benchmark.  The results from each example include error convergence plots and performance with time adaptation.  These examples prove it to be a viable and effective method for highly transient cases. More complex cases are intended to be applied to further test the method and its implementation.

>>>>

This report presents  the implementation of the Improved Quasi-Static method (IQS) in the Yak/Rattlesnake applications of the MOOSE framework.  In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.  The shape evaluation is very similar to a standard flux solve and is computed at large (macro) time steps. The amplitude is obtained by solving the PRKEs on a micro time scale. The PRKE parameters are regularly updated based on the shape changes.  Two versions of the IQS methodology have been implemented: (i) the standard approach, which leads to a system of nonlinear equations between the shape and the amplitude, and (ii) a predictor corrector approach that linearizes the system of equations. A time-adaptation technique has been developed to automatically select an appropriate macro time step size based on user-specified tolerances. The implementation has been verified and tested using manufactured solutions, a custom one-dimensional example, the TWIGL ramp benchmark, and the LRA benchmark.  The numerical results include error convergence plots and performance with time adaptation.  These examples prove it to be a viable and effective method for  transient neutronic simulations. More complex simulations (e.g., TREAT transients) are planned to perform coupled neutronics/heat conduction transient simulations within the MAMMOTH application.
